The zero--one laws of Kolmogorov and Hewitt--Savage in categorical probability

### 2019-11-14 categorical-zero-one.tex
Remove some superfluous uses of "clearly" &c

\begin{example}
    \label{setmulti}
    Let $\SetMulti$ be the category with objects ordinary sets, morphisms $f: X \to Y$ given by functions (i.e maps in the ordinary category of sets) $f: X \to \cP(Y)\setminus \emptyset$, and composition given by
    \[f\circ g(x) = \bigcup_{y \in g(x)} f(y)\]
    for $g: X \to Y, f: Y \to Z$.
    We think of the morphisms in $\SetMulti$ as \emph{multivalued functions}.
    See also \cite[Example~2.6]{markov_cats} for the analogous category of finite sets.
    The Cartesian product of sets gives it a symmetric monoidal structure, if we set \[(f \tensor g)(x_1,x_2) = f(x_1) \times g(x_2) \subset Y_1 \times Y_2,\] for $f:X_1 \to Y_1, g: X_2 \to Y_2$.
    This is a semicartesian monoidal structure, and we can give it a Markov category structure by setting
    $\cop{X}(x) = \{(x,x)\} \subset X \times X$.
    Then the deterministic morphisms are precisely those sending each point into a singleton set, and there is an obvious isomorphism of categories
    $\SetMulti_\detc \cong \sf{Set}$.
    Of course, $\sf{Set}$ actually has all infinite products, but they do not form Kolmogorov products, or even infinite tensor products in $\sf{SetMulti}$.
    To see this, consider the family $X_n = \{0,1\}$ for $n \in \bN$.
    A map $* \to \prod_{n\in \bN} \{0,1\} = X_\bN$ is just a subset of this infinite product.
    Given such a subset $U$, the finite marginals $* \to X_F$ for $F \subset \bN$ are the images of $U$ under the projection map.
    But these images do not determine $U$. For instance, consider the subset $U$ of those sequences containing at least one $1$.
    The image in $X_F$ is all of $X_F$ (since any finite sequence of $0$s and $1$s can be extended to one containing a $1$).
    But $U$ is not all of $X_\bN$.

@@

    It is also a surjection - given a family $(V_F) \in \lim H(\prod_{j \in J} X_j)$, simply set
    \[V = \{x \in \prod_{j\in J} X_j \mid  \pi_F(x) \in V_F \forall F\}\]
    This is an intersection of closed subsets, hence closed, and it is clearly sent to $(V_F)$.

@@

    It follows that $\prod_j X_j$ is an infinite tensor product in $\Kl(H)$ - the preservation condition is easily verified.
    The Kolmogorov product condition is similarly immediate to verify.
\end{proof}
\begin{corollary}
    Let $(X_j)_{j \in J}$ be a family of topological spaces, $Y$ a Hausdorff space, and let $f: \prod_j X_j \to Y$ be a continuous function
    which is independent of any finite prefix of the input.
    Then the map $f$ is constant.
\end{corollary}
\begin{corollary}
	Let $X$ be a topological space, $Y$ a Hausdorff space, $J$ an infinite set, and let $f: \prod_{j\in J} X \to Y$ be a continuous function which is invariant under any finite permutation of the inputs.
    Then $f$ is constant.
\end{corollary}
\begin{proof}[Proof of both]
	These results follow from \Cref{thm:kolmog,thm:hewsav} appliedto $\Kl(H)$.
	using the morphisms $p : I \to X_J$ corresponding to the trivial closed subset $X_J \subseteq X_J$.
    The conclusion is that the closure of the image, call it $A$, has the following property: $\cl(\{(a,a) \mid a \in A\}) = A \times A$ as subsets of $Y \times Y$.
    But since the diagonal is closed in $Y$, this clearly implies that $A$ must be a singleton.
\end{proof}

>>>>

\begin{example}
    \label{setmulti}
    Let $\SetMulti$ be the category with objects ordinary sets, morphisms $f: X \to Y$ given by functions (i.e maps in the ordinary category of sets) $f: X \to \cP(Y)\setminus \emptyset$, and composition given by
    \[f\circ g(x) = \bigcup_{y \in g(x)} f(y)\]
    for $g: X \to Y, f: Y \to Z$.
    We think of the morphisms in $\SetMulti$ as \emph{multivalued functions}.
    See also \cite[Example~2.6]{markov_cats} for the analogous category of finite sets.
    The Cartesian product of sets gives it a symmetric monoidal structure, if we set \[(f \tensor g)(x_1,x_2) = f(x_1) \times g(x_2) \subset Y_1 \times Y_2,\] for $f:X_1 \to Y_1, g: X_2 \to Y_2$.
    This is a semicartesian monoidal structure, and we can give it a Markov category structure by setting
    $\cop{X}(x) = \{(x,x)\} \subset X \times X$.
    Then the deterministic morphisms are precisely those sending each point into a singleton set, and there is an isomorphism of categories
    $\SetMulti_\detc \cong \sf{Set}$.
    Of course, $\sf{Set}$ actually has all infinite products, but they do not form Kolmogorov products, or even infinite tensor products in $\sf{SetMulti}$.
    To see this, consider the family $X_n = \{0,1\}$ for $n \in \bN$.
    A map $* \to \prod_{n\in \bN} \{0,1\} = X_\bN$ is just a subset of this infinite product.
    Given such a subset $U$, the finite marginals $* \to X_F$ for $F \subset \bN$ are the images of $U$ under the projection map.
    But these images do not determine $U$. For instance, consider the subset $U$ of those sequences containing at least one $1$.
    The image in $X_F$ is all of $X_F$ (since any finite sequence of $0$s and $1$s can be extended to one containing a $1$).
    But $U$ is not all of $X_\bN$.

@@

    It is also a surjection - given a family $(V_F) \in \lim H(\prod_{j \in J} X_j)$, simply set
    \[V = \{x \in \prod_{j\in J} X_j \mid  \pi_F(x) \in V_F \forall F\}\]
    This is an intersection of closed subsets, hence closed, and it is sent to $(V_F)$.

@@

    It follows that $\prod_j X_j$ is an infinite tensor product in $\Kl(H)$ - the preservation condition is easily verified.
    The Kolmogorov product condition is similarly immediate to verify.
\end{proof}
\begin{corollary}
    Let $(X_j)_{j \in J}$ be a family of topological spaces, $Y$ a Hausdorff space, and let $f: \prod_j X_j \to Y$ be a continuous function
    which is independent of any finite prefix of the input.
    Then the map $f$ is constant.
\end{corollary}
\begin{corollary}
	Let $X$ be a topological space, $Y$ a Hausdorff space, $J$ an infinite set, and let $f: \prod_{j\in J} X \to Y$ be a continuous function which is invariant under any finite permutation of the inputs.
    Then $f$ is constant.
\end{corollary}
\begin{proof}[Proof of both]
	These results follow from \Cref{thm:kolmog,thm:hewsav} appliedto $\Kl(H)$.
	using the morphisms $p : I \to X_J$ corresponding to the trivial closed subset $X_J \subseteq X_J$.
    The conclusion is that the closure of the image, call it $A$, has the following property: $\cl(\{(a,a) \mid a \in A\}) = A \times A$ as subsets of $Y \times Y$.
    But since the diagonal is closed in $Y$, this implies that $A$ must be a singleton.
\end{proof}
